---
title: "FA Assignment 1"
author: "Yunshuang Jiang"
date: "1/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#2 Download and analyze excess returns of S&P 500

Calculate continuous daily excess returns of SP500 (“^GSPC”) for the period from 1/1/2014 until 12/31/2014 using overnight Fed Funds rates as risk-free rates.
```{r message=FALSE}
library(quantmod)
getSymbols('^GSPC',from='2014-01-01',to='2014-12-31')
GSPC.return <- diff(log(GSPC$GSPC.Close))
library(readr)
FRB_H15 <- read_csv("FRB_H15.csv")[6:370,]
fedrate.return <- diff(log((as.numeric(FRB_H15$`Federal funds effective rate`))/360+1))
library(dplyr)
GSPC.return <- data.frame(Date=row.names(as.data.frame(GSPC.return)), Return=GSPC.return)
Fed.return <- data.frame(Date=FRB_H15$`Series Description`[-1], Return=fedrate.return)
join.data <- inner_join(x=GSPC.return,y=Fed.return, by="Date")
excess.return = (join.data$GSPC.Close-join.data$Return)*100
excessreturn = data.frame(Date = row.names(as.data.frame(GSPC.return)),
                            excess_return = excess.return)
excessreturn
```

#3 Download and analyze exchange rates

1) Find out how to download:
GBP/USD exchange rate, i.e. price of 1 British pound in US dollars, from Oanda using quantmode;
USD/JPY exchange rate, i.e. price of 1 US dollar in Japanese yen, from Oanda using quantmode.
```{r}
library(quantmod)
GBP.USD <- getSymbols.oanda("GBP/USD", auto.assign=FALSE)
USD.JPY <- getSymbols.oanda("USD/JPY", auto.assign=FALSE)
GBP.USD
USD.JPY
```


2) Calculate daily log returns of both exchange rates
```{r}
GBP.USD.logrtn<- diff(log(GBP.USD$GBP.USD))
USD.JPY.logrtn <- diff(log(USD.JPY$USD.JPY))
GBP.USD.logrtn
USD.JPY.logrtn
```


3) Calculate sample min, mean, sd, skewness, kurtosis, max of log returns for both exchange rates
```{r}
library(fBasics)
GBP.USD.stats <- basicStats(GBP.USD.logrtn)
GBP.USD.maxlogrtn <- max(GBP.USD.logrtn, na.rm=TRUE)
USD.JPY.stats <- basicStats(GBP.USD.logrtn)
USD.JPY.maxlogrtn <- max(USD.JPY.logrtn, na.rm=TRUE)
GBP.USD.stats
GBP.USD.maxlogrtn
USD.JPY.stats
USD.JPY.maxlogrtn
```


4) Test hypothesis H0rr:U=0 against alternative H0:U!=0
```{r}
t.test(as.data.frame(GBP.USD.logrtn)$GBP.USD)
t.test(as.data.frame(USD.JPY.logrtn)$USD.JPY)
```

