---
title: "FA Assignment 6"
author: "Yunshuang Jiang"
date: "2/15/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A stock index is currently at 810 and has volatility of 20%.\
The risk-free rate is 5% per year.\
Assume that dividend yield equals 1%.\
&nbsp;
&nbsp;

Price European 6-month call option with strike 800 using EuropeanOption() from RQuantLib.
Calculate the same premium manually using the formulas on the last slide of the lecture notes.\

```{r}
library(RQuantLib)

Option.Type = "call"
Underlying.Price = 810
Strike = 800
Option.Rate = 0.05
dividendYield = 0.01
volatility = 0.2
Expiry = 0.5

#Call without dividend
call= EuropeanOption(type=Option.Type, underlying=Underlying.Price,strike=Strike,
                     dividendYield=0,riskFreeRate=Option.Rate, maturity=Expiry, volatility=volatility)
call

#Call without dividend manually
d1 = (log(Underlying.Price/Strike) + (Option.Rate+(volatility^2)/2)*Expiry)/ (volatility*sqrt(Expiry))
d2 = d1 - (volatility*sqrt(Expiry))
call.manual = Underlying.Price*pnorm(d1) - Strike*(exp(-Option.Rate*Expiry))*pnorm(d2)
call.manual  

```
&nbsp;
&nbsp;

Think how dividend yield should affect option price.\
Comment: With dividend yield, the call price should decrease.\

Experiment with the function EuropeanOption() with zero or non-zero dividend yield and find how the Black-Scholes formula on slide 17 should be modified for dividend yield.\

```{r}
#Call with dividend
call.withdiv= EuropeanOption(type=Option.Type, underlying=Underlying.Price,strike=Strike, dividendYield=dividendYield,riskFreeRate=Option.Rate, maturity=Expiry, volatility=volatility)
call.withdiv

#Call with dividend manually
d1.div = (log(Underlying.Price/Strike) + (Option.Rate-dividendYield+(volatility^2)/2)*Expiry)/ (volatility*sqrt(Expiry))
d2.div = d1.div - (volatility*sqrt(Expiry))
call.div.manual = (Underlying.Price*exp(-dividendYield*Expiry))*pnorm(d1.div) - Strike*(exp(-Option.Rate*Expiry))*pnorm(d2.div)
call.div.manual
```

Black-Scholes formula with dividend rate q.\
$$d1 = \frac{(ln(\frac{S_{0}}{K})+(R-q+ \frac{\sigma^2}{2}) \times T)}{\sigma \times \sqrt{T}}$$ 
$$d2 = d1 - \sigma \times \sqrt{T}$$  
$$C = S_{0}\times e^{qT} \times \mathcal{N}(d1) -Ke^{-RT}\times \mathcal{N}(d2)$$

Calculate the premium of put option with the same expiry and same strike using put-call parity.\

```{r}
#put without dividend
put = call$value + exp(-Option.Rate*Expiry)*Strike - Underlying.Price
put
```

```{r}
#put with dividend
put.withdiv = call.withdiv$value + exp(-Option.Rate*Expiry)*Strike*(1+dividendYield) - Underlying.Price
put.withdiv
```

