---
title: "FA Assignment 4"
author: "Yunshuang Jiang"
date: "2/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

7. Consider the quarterly earnings per share of the Johnson & Johnson from the first quarter of 1992 to the second quarter of 2011. The data are in the file q-jnj-earns-9211.txt and are obtained from the First Call Historical Database of Thomson Reuters. Take log transformation of the data if necessary. Build a time series model for the data. Perform model checking to assess the adequacy of the fitted model. Write down the model. Refit the model using data from 1992 to 2008. Perform 1-step to 10-step ahead forecasts of the quarterly earnings and obtain a forecast plot.

```{r}
#load data
jj <- read.csv("~/Downloads/assignments_Fonancial Analytics (32001)_Lecture 4_q-jnj-earns-9211.txt", sep="")

#initial graphs 
plot(jj$earns,type='l', ylab='growth') 

#the plot suggests seasonal patterns, we try taking the log and difference of the earnings
jjlog=log(jj$earns)
djjlog=diff(jjlog)
djjlog4=diff(jjlog,4)
ddjjlog=diff(djjlog4)
par(mfcol=c(2,2))
acf(jjlog,lag=20)
acf(djjlog,lag=20)
acf(djjlog4,lag=20)
acf(ddjjlog,lag=20)

c1=c("2","3","4","1")
c2=c("1","2","3","4")
par(mfcol=c(3,1))
plot(djjlog,ylab='diff',type='l')
points(djjlog,pch=c1,cex=0.7)
plot(djjlog4,ylab='sea-diff',type='l')
points(djjlog4,pch=c2,cex=0.7)
plot(ddjjlog,ylab='dd',type='l')
points(ddjjlog,pch=c1,cex=0.7) 

```

```{r}
#build time series model

m1=arima(jjlog,order=c(0,1,2),seasonal=list(order=c(0,1,1),period=4))
m2=arima(jjlog,order=c(0,1,1),seasonal=list(order=c(1,1,0),period=4))
m3=arima(jjlog,order=c(2,1,0),seasonal=list(order=c(0,1,1),period=4))
m4=arima(jjlog,order=c(1,1,0),seasonal=list(order=c(1,1,0),period=4))

aic <- c("m1" =m1$aic, "m2"=m2$aic, "m3"=m3$aic, "m4"=m4$aic)
aic

#Based on AIC, we select model 1 as our final model. 
m1

#model check 
tsdiag(m1,gof=20) 
Box.test(m1$residuals,lag=12,type='Ljung')
pp=1-pchisq(2.8062,11)
pp

#Conclusion: Since pvalue = 0.993047 > alpha = 0.05, 
#so we cannot reject the hypothesis that all autocorrelations are equal to zero.
```


Fitted Model:
$$(1-B) \times (1-B^4) \times X_t = (1 -0.1796B^4)(1 -0.2713B  -0.1693B^2) \times Z_t$$ 


```{r}
#refit the model using data from 1992-2008

#1)subset the data
jjnew <- jj[1:68,]
jjnewlog <- log(jjnew$earns)

#2)refit the model
m1new=arima(jjnewlog,order=c(0,1,2),seasonal=list(order=c(0,1,1),period=4))


#3)model check 
tsdiag(m1new,gof=20) 
Box.test(m1new$residuals,lag=12,type='Ljung')
pp=1-pchisq(2.6892,11)
pp

```

```{r}
#perform 1-step to 10-step ahead forecast of quarterly earnings
prd<-predict(m1new,10)

#forecast plot
pred=prd$pred
se=prd$se
ko=jj$earns # actual observations
fore=exp(pred+se^2/2) #point forecasts, delogged
v1=exp(2*pred+se^2)*(exp(se^2)-1)
s1=sqrt(v1) # std of the forecast error
eps=ko[50:78]

tdx=(c(1:length(eps)))/4+1992
upp=c(ko[68],fore+2*s1) # upper band (+2*std)
low=c(ko[68],fore-2*s1) # lower band (-2*std)
min(low,eps)
max(upp,eps)

plot(tdx,eps,xlab='year',ylab='earnings',type='l',ylim=c(0.50,2.5))
points(tdx[(length(eps)-9):length(eps)],fore,pch='*')
lines(tdx[(length(eps)-10):length(eps)],upp,lty=2)
lines(tdx[(length(eps)-10):length(eps)],low,lty=2)
points(tdx[(length(eps)-9):length(eps)],ko[69:78],pch='o',cex=0.7)

```

