---
title: "FA Assignment 8"
author: "Yunshuang Jiang"
date: "3/5/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Exercise 1 on page 239.\
Answer questions using 5% significance level in tests and 10 lags of serial correlations for return series.\

Consider daily returns of ETF SPDR S&P 500 from file d-spy-0111.txt.\
Transform the simple returns into log-returns\

```{r}
dat <- read.csv("~/Downloads/assignments_Fonancial Analytics (32001)_Lecture 8_d-spy-0111.txt", sep="")
dat$rtn = log(dat$rtn+1)
rtn=ts(dat$rtn,frequency=250,start=c(2001,9))
```


a). Is the expected log-return zero? Are there any serial correlations in the log returns? Is there ARCH effect in the log-returns?\

```{r}
plot(rtn,type='l',xlab='year',ylab='rtn')
t.test(dat$rtn)
#the p-value is not small enough to reject the null hypothesis, 
#hence we conclude that the true mean of the log return is equal to 0.

Box.test(dat$rtn,lag=10,type='Ljung')
Box.test(abs(dat$rtn),lag=10,type='Ljung')
#the p-values for both the log return and the absolute log returns are 
#small enough to reject the null hypothesis, hence we conclude that 
#the serial correlations are not zero. 

par(mfcol=c(2,1))
acf(dat$rtn,lag=10) 
acf(abs(dat$rtn),lag=10) 
#While ACF of returns is not significantly different from zero for most
#lags, the ACF of absolute log returns is outside the bounds.
```

```{r}
## ARCH test
y=dat$rtn-mean(dat$rtn)
Box.test(y^2,lag=10,type='Ljung')
#According to Box-Ljung test the zero correlations hypothesis is rejected.
#This is an indication of ARCH effect.

#ARCH test
archTest <- function(rtn,m=10){
  # Perform Lagrange Multiplier Test for ARCH effect of a time series
  # rtn: time series
  # m: selected AR order
  #
  y=(rtn-mean(rtn))^2
  T=length(rtn)
  atsq=y[(m+1):T]
  x=matrix(0,(T-m),m)
  for (i in 1:m){
    x[,i]=y[(m+1-i):(T-i)]
  }
  md=lm(atsq~x)
  summary(md)
}

archTest(y,10)
#The p-value is small, hence the output confirms ARCH effect.
```


b). Fit Gaussian ARMA-GARCH model for the log-return series. Perform model checking. Obtain the QQ-plot of the standardized residuals. Write down the fitted model. [Hint: use GARCH(2,1)].\

```{r}
acf(dat$rtn^2)
pacf(dat$rtn^2)
suppressWarnings(library(fGarch))
garch1=garchFit(~1+garch(2,1),data=dat$rtn,trace=F)
summary(garch1)
```

Fitted Model:\

$$ r_t = 0.00057243 + \sigma_t   \epsilon_t $$
$$ \sigma_t^2 = 0.0000023226 + 0.0019896 a_{t-1}^2 + 0.11165 a_{t-2}^2 + 0.87049 \sigma_{t-1}^2$$

```{r}
#Analyze the residuals
resi=residuals(garch1,standardize=T)
tdx=c(1:2535)/250+2001
par(mfcol=c(3,1))
plot(tdx,resi,xlab='year',ylab='stand-resi',type='l')
acf(resi,lag=10)
pacf(resi^2,lag=10)

#The ACF plot shows that standardized residuals do not have autocorrelation.
#PACF for squared residuals shows there might be significant correlation at higher lags (log=10).

qqnorm(resi, pch = 1, frame = FALSE)
qqline(resi, col = "steelblue", lwd = 2)
#The qq plot looks normal from -2 to 2 with some points at both tails are slightly off. 
#We conclude that the residuals is distributed generally normal. 
```


c). Build an ARMA-GARCH model with Student t innovations for the log-return series. Perform model checking and write down the fitted model.\

```{r}
garch2=garchFit(~1+garch(2,1),data=dat$rtn,trace=F,cond.dist="std")
summary(garch2)

#Analyze the residuals
tresi=residuals(garch2,standardize=T)
tdx=c(1:2535)/250+2001
par(mfcol=c(3,1))
plot(tdx,tresi,xlab='year',ylab='stand-resi',type='l')
acf(tresi,lag=10)
pacf(tresi^2,lag=10)

#The ACF plot shows that standardized residuals do not have autocorrelation.
#PACF for squared residuals shows no significant correlation.

qqnorm(tresi, pch = 1, frame = FALSE)
qqline(tresi, col = "steelblue", lwd = 2)
#The qq plot looks normal from -1 to 3 with some points at both left tails are slightly off. 
#We conclude that the residuals is distributed generally normal.
```

Fitted Model:\

$$ r_t = 0.00072454 + \sigma_t   \epsilon_t $$
$$ \sigma_t^2 = 0.0000015793 + 0.0061117 a_{t-1}^2 + 0.11435 a_{t-2}^2 + 0.87418 \sigma_{t-1}^2$$